Practical Parametricity for GADTs

### 2020-09-17 20-gadts/inf-rules.tex
actions of interps of Lans defined on objects and IEL proved(?); actions of interps on morphisms still to be defined

{\color{red} We should ask what, intuitively, it means to have
  parametricity for a calculus with GADTs. In particular it should
  mean that any polymorphic function that takes a GADT as input should
  operate uniformly over all type instances of the GADT. That means
  that the GADT must itself be ``parametric'', i.e., uniform in its
  type indices. But what does this mean? For ADTs and nested types it
  means that the datatype elements are constructed in exactly the same
  ways at each type instance. For GADTs, it means that as well, but
  here we also have to take into account which instances are actually
  inhabited, since not all type instances of a GADT are inhabited.
  [Of course we did not have to do this for ADTs and nested types
    because {\em all} instances of ADTs and nested types are
    inhabited.] So, intuitively, we consider a GADT to be
  ``parametric'' if 1) the corresponding instances of the GADT are
  inhabited for for every instantiation of its type parameters
  $\ol\alpha$, and 2) the objects for each instantiation of the
  $\alpha$s are constructed in exactly the same way. That is, which
  instances are inhabited, as well as how the data in each of those
  instances are constructed, is type independent. We capture this in
  our calculus by requiring that, for $\Gamma;\Phi \vdash
  (\Lan^{\ol{\alpha^0}}_{\ol K} F)\, \ol A$, we have that
  $\emptyset;\ol\beta \vdash K$, and that $K$ does not involve any
  constants. (The latter ensures that a GADT with a constructor of
  type $\alpha \to Int$ is not parametric, which makes sense since
  there will be more ways to construct Int instances of the GADT than
  ways to construct its other type instances.)  Further, to ensure
  that the categorical Lans interpreting $\Lan$-constructs are
  $\omega$-cocontinuous we also require that $K$ is a polynomial
  functor in the $\alpha$s that does not involve any constant
  functors. This means that our constructors have types of the form
  $F\ol\beta \to F(K\ol\beta)$ with no constants in $K$, so that
  exactly the corresponding instances of the GADT are inhabited for
  every instantiation of the $\alpha$s, and the objects for each
  instantiation of the $\alpha$s are constructed in exactly the same
  way. That is, the inhabited instances and the construction of the
  data in those instances are type independent.}

@@

\[\begin{array}{c}
\AXC{$\Gamma;\Phi \vdash F$}
\AXC{$\Gamma;\Phi  \vdash G$}
\BIC{$\Gamma;\emptyset \vdash \Nat^\Phi F \,G$}
\DisplayProof
\\[4ex]
\AXC{$\phi^k \in \Gamma \cup \Phi$}
\AXC{$\quad\quad\ol{\Gamma;\Phi \vdash F}$}
\BIC{$\Gamma;\Phi \vdash \phi^k \ol{F}$}
\DisplayProof
\\[3ex]
\AXC{$\Gamma;\ol{\gamma^0},\ol{\alpha^0},\phi^k \vdash F$}
\AXC{$\quad\quad\ol{\Gamma;\Phi,\ol{\gamma^0} \vdash G}$}
\BIC{$\Gamma;\Phi,\ol{\gamma^0} \vdash (\mu \phi^k.\lambda
  \ol{\alpha^0}. \,F)\,\ol{G}$} 
\DisplayProof
\\[5ex]
\AXC{$\Gamma; \Phi,\ol\alpha^0 \vdash F$}
\AXC{$\ol{{\color{blue} \emptyset};\ol\alpha^0 \vdash K}$}
\AXC{$\ol{\Gamma;\Phi \vdash A}$}
\TIC{$\Gamma;\Phi \vdash (\Lan^{\ol{\alpha^0}}_{\ol K} F)\, \ol A$}
\DisplayProof
\end{array}\]
\end{dfn}

@@

\[\begin{array}{c}
\AXC{$\Gamma; \Phi \vdash F$}
\AXC{$\Gamma; \Phi \vdash G$}
\AXC{$\Gamma; \Phi \,|\, \Delta, x : F \vdash t: G$} 
\TIC{$\Gamma; \emptyset
  \,|\, \Delta \vdash L_\Phi x.t : \Nat^\Phi \,F \,G$}
\DisplayProof
\\\\
\AXC{$\ol{\Gamma;\Phi,\ol\beta \vdash K}$}
\AXC{$\Gamma; \emptyset
  \,|\, \Delta \vdash t : \Nat^{\ol\psi} \,F \,G$}
\AXC{$\Gamma;\Phi \,|\, \Delta \vdash s: F[\overline{\psi :=_{\ol\beta} K}]$}
\TIC{$\Gamma;\Phi\,|\, \Delta \vdash t_{\ol K} s:
  G[\overline{\psi :=_{\ol\beta} K}]$}
\DisplayProof
\\\\
\AXC{$\Gamma; \Phi, \ol{\phi} \vdash H$}
\AXC{$\ol{\Gamma; \Phi,\ol{\beta} \vdash F}$}
\AXC{$\ol{\Gamma; \Phi,\ol{\beta} \vdash
    G}$}
\TIC{$\Gamma; \emptyset
  ~|~\emptyset
  \vdash \map^{\ol{F},\ol{G}}_H :
  \Nat^\emptyset\;(\ol{\Nat^{\Phi,\ol{\beta}}\,F\,G})\;
  (\Nat^\Phi\,H[\ol{\phi :=_{\ol{\beta}} F}]\;H[\ol{\phi
      :=_{\ol{\beta}} G}])$} 
\DisplayProof
\\\\
\AXC{$\Gamma; \Phi, \phi, \ol{\alpha} \vdash H$}
\UIC{$\Gamma; \emptyset  \,|\, \emptyset \vdash \tin_H :
  \Nat^{\Phi,\ol{\beta}} H[\phi :=_{\ol{\beta}} (\mu
    \phi.\lambda \ol{\alpha}.H)\ol{\beta}][\ol{\alpha := \beta}]\,(\mu
  \phi.\lambda \ol{\alpha}.H)\ol{\beta}$}
\DisplayProof
\\\\
\AXC{$\Gamma; \phi,\Phi,\ol{\alpha} \vdash H$}
\AXC{$\Gamma; \Phi, \ol{\beta} \vdash F$}
\BIC{$\Gamma; \emptyset  \,|\, \emptyset \vdash \fold^F_H :
  \Nat^\emptyset\; (\Nat^{\Phi,\ol{\beta}}\,H[\phi
    :=_{\ol{\beta}} F][\ol{\alpha := \beta}]\,F)\; (\Nat^{\Phi,\ol{\beta}}
    \,(\mu \phi.\lambda \ol{\alpha}.H)\ol{\beta}\,F)$}
\DisplayProof
\\\\
\AXC{$\Gamma; \Phi,\ol{\alpha}\vdash F$}
\AXC{$\ol{{\color{blue} \emptyset}; \ol{\alpha} \vdash K}$}
\AXC{$\ol{\Gamma;\Phi \vdash A} \hspace*{0.3in} \Gamma;\Phi~|~\Delta \vdash t : F[\ol{\alpha :=A}]$}
\TIC{$\Gamma;\Phi~|~\Delta \vdash \int_{\ol K,F} t :
  (\Lan^{\ol\alpha}_{\ol K}\,  F)\,\ol{K[\ol{\alpha := A}]}$}  
\DisplayProof
\\\\
\AXC{$\Gamma;\emptyset~|~\Delta \vdash \eta : \Nat^{\Phi,\ol\alpha}\,
  F\;G[\ol{\beta := K}]$}
\AXC{$\ol{\Gamma;\Phi \vdash B}$}
\AXC{$\Gamma;\Phi~|~\Delta \vdash t : (\Lan^{\ol\alpha}_{\ol K} F)\,\ol B $}
\TIC{$\Gamma;\Phi~|~\Delta \vdash \partial^{G, \ol K}_F \eta\,t :
 G[\ol{\beta := B}]$}
\DisplayProof
\end{array}\]
\end{dfn}

@@

To accommodate GADTs, we will need to transition Reynolds' approach
from a $\set$-based semantics to a semantics based on
$\omega$-complete partial orders. We denote the category of
$\omega$-complete partial orders ($\oCPO$s) and their {\color{blue}
  monotone?} sup-preserving morphisms by $\omega$CPO. The underlying
set of an $\oCPO$ $A$ is denoted $|A|$. The category $\oCPOR$ of
$\oCPO$ relations has as its objects triples $(A,B,R)$, where $A, B:
\oCPO$; $R: \rel(|A|, |B|)$; {\color{blue} $(a,b) \in R$, $a \leq a'$,
  and $b \leq b'$ imply $(a',b') \in R$; so next part is redundant?}
and $(\bigvee_{i < \omega} a_i, \bigvee_{i < \omega} b_i) \in R$
whenever $(a_i)_{i < \omega}$ and $(b_i)_{i < \omega}$ are chains in
$A$ and $B$, respectively, such that $(a_i,b_i) \in R$ for all
$i$. The category $\oCPOR$ has as its morphisms from $(A,B,R)$ to
$(A',B',R')$ pairs $(f : A \to A',g : B \to B')$ of morphisms in
$\oCPO$ such that $(f\,a, g\,b) \in R'$ whenever $(a,b) \in R$. We
note that if $(f,g) : (A,B,R) \to (A',B',R')$ and $(a_i)_{i < \omega}$
and $(b_i)_{i < \omega}$ are chains in $A$ and $B$, respectively, then
$ (f \, (\bigvee_{i < \omega} a_i), g \, (\bigvee_{i < \omega} b_i)) =
(\bigvee_{i < \omega} (f a_i), \bigvee_{i < \omega} (g b_i)) \in R'$
necessarily holds.  We write $R : \oCPOR(A,B)$ in place of $(A,B,R):
\oCPOR$ when convenient. If $R : \oCPOR(A,B)$ then we write $\pi_1 R$
and $\pi_2 R$ for the {\em domain} $A$ of $R$ and the {\em codomain}
$B$ of $R$, respectively. If $A : \oCPO$, then we write $\Eq_A =
(A,A,\Eq_{|A|})$ for the {\em equality relation} on $A$.

@@

The object interpretations of the types in our calculus will be
$\omega_1$-cocontinuous functors between categories of
$\omega_1$-cocontinuous functors on categories constructed from the
locally $\omega_1$-presentable category $\oCPO$.  We therefore begin
by recording some important facts about locally $\omega_1$-presentable
categories and functors on them, and verifying verify the properties
needed to interpret our syntax.

@@

\new{Perhaps have as preliminaries to entire paper. Do everything for
  $\lambda CPOs$? Define these; investigate their properties.}

@@

A poset $\mcD = (D,\leq)$ is \emph{$\omega_1$-directed} if every
countable subset of $D$ has a supremum. When $\mcD$ is considered
as a category, we write $d \in \mcD$ to indicate that $d$ is an object
of $\mcD$ (i.e., $d \in D$). An {\em $\omega_1$-directed colimit} in a
category $\mcC$ is a colimit of a diagram $F : {\mathcal D} \to \mcC$,
where $\mathcal D$ is an $\omega_1$-directed poset. A category $\mcC$
is {\em $\omega_1$-cocomplete} if it has all $\omega_1$-directed
colimits; a {\em cocomplete} category is one that has all colimits.

If $\mcA$ and $\mcC$ are $\omega_1$-cocomplete, then a functor $F :
\mcA \to \mcC$ is {\em $\omega_1$-cocontinuous} if it preserves
$\omega_1$-directed colimits. We write $[\A,\C]_{\omega_1}$ for the
category of $\omega_1$-cocontinuous functors from $\A$ to $\C$, and
$\C^\A$ for the category of {\em all} functors from $\A$ to $\C$.
Since (co)limits of functors are computed pointwise, $C^\A$ has all
(co)limits that $\C$ has, and (co)limits of (co)continuous functors
are again (co)continuous. It follows that $[\A,\C]_{(\omega_1)}$ is
($\omega_1$-)(co)complete whenever $\C$ is. In particular, if $F :
[\C, \C]_{(\omega_1)}$ for a(n) ($\omega_1$-)cocomplete category $\C$,
then we can define $F^0 = \Id$, $F^{\alpha + 1} = F \circ F^\alpha$ at
successor ordinals, and $F^{\bigcup_{\beta < \alpha} \beta} =
\colim{\beta < \alpha} F^\beta$ at limit ordinals.

If $\mcA$ is locally small, then an object $A$ of $\mcA$ is {\em
  $\omega_1$-presentable} if the functor $\mathsf{Hom}_\mcA(A, -) :
\mcA \to \Set$ preserves $\omega_1$-directed colimits, i.e., if for
every $\omega_1$-directed poset $\mathcal D$ and every functor $F :
{\mathcal D} \to \A$, there is a canonical isomorphism $\colim{d \in
  \mcD}{\mathsf{Hom}_\mcA(A,Fd)} \simeq \mathsf{Hom}_{\mcA}(A,
\colim{d \in \mcD}{Fd})$. A locally small category $\mcA$ is {\em
  $\omega_1$-accessible} if it is $\omega_1$-cocomplete and has a set
$\mcA_0$ of $\omega_1$-presentable objects such that every object is
an $\omega_1$-directed colimit of objects in $\mcA_0$; a locally small
category is {\em locally $\omega_1$-presentable} if it is
$\omega_1$-accessible and cocomplete.

The category $\oCPO$ is locally $\omega_1$-presentable (but not
locally finitely presentable); see Examples~1.18(2)
of~\cite{ar94}). Its $\omega_1$-presentable objects are precisely the
$\omega$CPOs cardinality less than $\omega_1$, i.e., the countable
$\omega$CPOs. In the next subsection we will interpret type variables
in $\tvars^k \cup \fvars^k$ as elements of
$[\oCPO^k,\oCPO]_{\omega_1}$; the following special cases of standard
results (see, e.g.,~\cite{ar94}) will therefore be critical to
deducing important properties of our object interpretations of types:
\begin{prop}\label{prop:prelims}
\begin{enumerate}
\item If $\C_1,...,\C_n$ are locally $\omega_1$-presentable categories
  then so is $\C_1 \times ... \times \C_n$. Moreover, the presentable
  objects of $\C_1 \times ... \times \C_n$ are exactly the tuples of
  the form $(P_1,...,P_n)$, where, for each $i = 1,...,n$, the object
  $P_i$ is presentable in $\C_i$.
\item If $\mathcal{A}$ is $\omega_1$-accessible and $\mathcal{C}$ is
  $\lambda$-cocomplete, then the category $[\mathcal{A},
  \mathcal{C}]_{\omega_1}$ is naturally equivalent to the category
  $\mathcal{C}^{\mathcal{A}_0}$.
\item If $\mathcal{C}$ is locally $\omega_1$-presentable and
  $\mathcal{A}_0$ is essentially small, then
  $\mathcal{C}^{\mathcal{A}_0}$ is locally $\omega_1$-presentable.
\end{enumerate}
\end{prop}
\noindent
Together, the statements in Proposition~\ref{prop:prelims} give that
if $\mathcal{A}$ and $\mathcal{C}$ are locally $\omega_1$-presentable,
then $[\mathcal{A}, \mathcal{C}]_{\omega_1}$ is naturally equivalent
to $\mathcal{C}^{\mathcal{A}_0}$, and thus is
$\omega_1$-presentable. Thus, for all $k_1,...,k_n \in \nat^n$,
$[\mathcal{A}^{k_1}, \mathcal{C}]_{\omega_1} \times ... \times
[\mathcal{A}^{k_n}, \mathcal{C}]_{\omega_1}$ is locally
$\omega_1$-presentable, and therefore $[[\mathcal{A}^{k_1},
    \mathcal{C}]_{\omega_1} \times ... \times [\mathcal{A}^{k_n},
    \mathcal{C}]_{\omega_1}, \mathcal{C}]_{\omega_1}$ is as
well. Taking both $\A$ and $C$ to be $\oCPO$ --- as we will to ensure
that the fixpoints interpreting $\mu$-types in $\oCPO$ exist --- we
have
\begin{prop}\label{prop:nat-interps-sets}
  For all $k_1,...,k_n \in \nat^n$,
  \[[[\oCPO^{k_1}, \oCPO]_{\omega_1}
  \times ... \times [\oCPO^{k_n}, \oCPO]_{\omega_1},
  \oCPO]_{\omega_1}\] is locally $\omega_1$-presentable.
\end{prop}

@@

\begin{dfn}\label{def:set-env}
A {\em $\oCPO$ environment} maps each type variable in $\tvars^k \cup
\fvars^k$ to an element of $[\oCPO^k,\oCPO]_{\omega_1}$.  A morphism
$f : \rho \to \rho'$ for set environments $\rho$ and $\rho'$ with
$\rho|_\tvars = \rho'|_\tvars$ maps each type constructor variable
$\psi^k \in \tvars$ to the identity natural transformation on $\rho
\psi^k = \rho'\psi^k$ and each functorial variable $\phi^k \in \fvars$
to a natural transformation from the $k$-ary functor $\rho \phi^k$ on
$\oCPO$ to the $k$-ary functor $\rho' \phi^k$ on $\oCPO$.  Composition
of morphisms on set environments is given componentwise, with the
identity morphism mapping each set environment to itself. This gives a
category of set environments and morphisms between them, which we
denote $\oCPOenv$.
\end{dfn}

>>>>

{\color{red} We should ask what, intuitively, it means to have
  parametricity for a calculus with GADTs. In particular it should
  mean that any polymorphic function that takes a GADT as input should
  operate uniformly over all type instances of the GADT. That means
  that the GADT must itself be ``parametric'', i.e., uniform in its
  type indices. But what does this mean? For ADTs and nested types it
  means that the datatype elements are constructed in exactly the same
  ways at each type instance. For GADTs, it means that as well, but
  here we also have to take into account which instances are actually
  inhabited, since not all type instances of a GADT are inhabited.
  [Of course we did not have to do this for ADTs and nested types
    because {\em all} instances of ADTs and nested types are
    inhabited.] So, intuitively, we consider a GADT to be
  ``parametric'' if 1) the corresponding instances of the GADT are
  inhabited for for every instantiation of its type parameters
  $\ol\alpha$, and 2) the objects for each instantiation of the
  $\alpha$s are constructed in exactly the same way. That is, which
  instances are inhabited, as well as how the data in each of those
  instances are constructed, is type independent. We capture this in
  our calculus by requiring that, for $\Gamma;\Phi \vdash
  (\Lan^{\ol{\alpha^0}}_{\ol K} F)\, \ol A$, we have that
  $\emptyset;\ol\beta \vdash K$, and that $K$ does not involve any
  constants. (The latter ensures that a GADT with a constructor of
  type $\alpha \to Int$ is not parametric, which makes sense since
  there will be more ways to construct Int instances of the GADT than
  ways to construct its other type instances.) Further, to ensure that
  the categorical Lans interpreting $\Lan$-constructs are
  $\omega$-cocontinuous we also require that $K$ is a polynomial
  functor in the $\alpha$s (that necessarily will not involve any
  constant functors). This means that our constructors have types of
  the form $F\ol\beta \to F(K\ol\beta)$ with no constants in $K$, so
  that exactly the corresponding instances of the GADT are inhabited
  for every instantiation of the $\alpha$s, and the objects for each
  instantiation of the $\alpha$s are constructed in exactly the same
  way. That is, the inhabited instances and the construction of the
  data in those instances are type independent.}

@@

\[\begin{array}{c}
\AXC{$\Gamma;\Phi \vdash F$}
\AXC{$\Gamma;\Phi  \vdash G$}
\BIC{$\Gamma;\emptyset \vdash \Nat^\Phi F \,G$}
\DisplayProof
\\[4ex]
\AXC{$\phi^k \in \Gamma \cup \Phi$}
\AXC{$\quad\quad\ol{\Gamma;\Phi \vdash F}$}
\BIC{$\Gamma;\Phi \vdash \phi^k \ol{F}$}
\DisplayProof
\\[3ex]
\AXC{$\Gamma;\ol{\gamma^0},\ol{\alpha^0},\phi^k \vdash F$}
\AXC{$\quad\quad\ol{\Gamma;\Phi,\ol{\gamma^0} \vdash G}$}
\BIC{$\Gamma;\Phi,\ol{\gamma^0} \vdash (\mu \phi^k.\lambda
  \ol{\alpha^0}. \,F)\,\ol{G}$} 
\DisplayProof
\\[5ex]
\AXC{$\Gamma; \Phi,\ol\alpha^0 \vdash F$}
\AXC{$\ol{\Gamma;\ol\alpha^0 \vdash K}$}
\AXC{$\ol{\Gamma;\Phi \vdash A}$}
\TIC{$\Gamma;\Phi \vdash (\Lan^{\ol{\alpha^0}}_{\ol K} F)\, \ol A$}
\DisplayProof
\end{array}\]
\end{dfn}

@@

\[\begin{array}{c}
\AXC{$\Gamma; \Phi \vdash F$}
\AXC{$\Gamma; \Phi \vdash G$}
\AXC{$\Gamma; \Phi \,|\, \Delta, x : F \vdash t: G$} 
\TIC{$\Gamma; \emptyset
  \,|\, \Delta \vdash L_\Phi x.t : \Nat^\Phi \,F \,G$}
\DisplayProof
\\\\
\AXC{$\ol{\Gamma;\Phi,\ol\beta \vdash K}$}
\AXC{$\Gamma; \emptyset
  \,|\, \Delta \vdash t : \Nat^{\ol\psi} \,F \,G$}
\AXC{$\Gamma;\Phi \,|\, \Delta \vdash s: F[\overline{\psi :=_{\ol\beta} K}]$}
\TIC{$\Gamma;\Phi\,|\, \Delta \vdash t_{\ol K} s:
  G[\overline{\psi :=_{\ol\beta} K}]$}
\DisplayProof
\\\\
\AXC{$\Gamma; \Phi, \ol{\phi} \vdash H$}
\AXC{$\ol{\Gamma; \Phi,\ol{\beta} \vdash F}$}
\AXC{$\ol{\Gamma; \Phi,\ol{\beta} \vdash
    G}$}
\TIC{$\Gamma; \emptyset
  ~|~\emptyset
  \vdash \map^{\ol{F},\ol{G}}_H :
  \Nat^\emptyset\;(\ol{\Nat^{\Phi,\ol{\beta}}\,F\,G})\;
  (\Nat^\Phi\,H[\ol{\phi :=_{\ol{\beta}} F}]\;H[\ol{\phi
      :=_{\ol{\beta}} G}])$} 
\DisplayProof
\\\\
\AXC{$\Gamma; \Phi, \phi, \ol{\alpha} \vdash H$}
\UIC{$\Gamma; \emptyset  \,|\, \emptyset \vdash \tin_H :
  \Nat^{\Phi,\ol{\beta}} H[\phi :=_{\ol{\beta}} (\mu
    \phi.\lambda \ol{\alpha}.H)\ol{\beta}][\ol{\alpha := \beta}]\,(\mu
  \phi.\lambda \ol{\alpha}.H)\ol{\beta}$}
\DisplayProof
\\\\
\AXC{$\Gamma; \phi,\Phi,\ol{\alpha} \vdash H$}
\AXC{$\Gamma; \Phi, \ol{\beta} \vdash F$}
\BIC{$\Gamma; \emptyset  \,|\, \emptyset \vdash \fold^F_H :
  \Nat^\emptyset\; (\Nat^{\Phi,\ol{\beta}}\,H[\phi
    :=_{\ol{\beta}} F][\ol{\alpha := \beta}]\,F)\; (\Nat^{\Phi,\ol{\beta}}
    \,(\mu \phi.\lambda \ol{\alpha}.H)\ol{\beta}\,F)$}
\DisplayProof
\\\\
\AXC{$\Gamma; \Phi,\ol{\alpha}\vdash F$}
\AXC{$\ol{\Gamma; \ol{\alpha} \vdash K}$}
\AXC{$\ol{\Gamma;\Phi \vdash A} \hspace*{0.3in} \Gamma;\Phi~|~\Delta \vdash t : F[\ol{\alpha :=A}]$}
\TIC{$\Gamma;\Phi~|~\Delta \vdash \int_{\ol K,F} t :
  (\Lan^{\ol\alpha}_{\ol K}\,  F)\,\ol{K[\ol{\alpha := A}]}$}  
\DisplayProof
\\\\
\AXC{$\Gamma;\emptyset~|~\Delta \vdash \eta : \Nat^{\Phi,\ol\alpha}\,
  F\;G[\ol{\beta := K}]$}
\AXC{$\ol{\Gamma;\Phi \vdash B}$}
\AXC{$\Gamma;\Phi~|~\Delta \vdash t : (\Lan^{\ol\alpha}_{\ol K} F)\,\ol B $}
\TIC{$\Gamma;\Phi~|~\Delta \vdash \partial^{G, \ol K}_F \eta\,t :
 G[\ol{\beta := B}]$}
\DisplayProof
\end{array}\]
\end{dfn}

@@

{\color{red} NOT NEEDED ANYMORE??
To accommodate GADTs, we will need to transition Reynolds' approach
from a $\set$-based semantics to a semantics based on
$\omega$-complete partial orders. We denote the category of
$\omega$-complete partial orders ($\oCPO$s) and their {\color{blue}
  monotone?} sup-preserving morphisms by $\omega$CPO. The underlying
set of an $\oCPO$ $A$ is denoted $|A|$. The category $\oCPOR$ of
$\oCPO$ relations has as its objects triples $(A,B,R)$, where $A, B:
\oCPO$; $R: \rel(|A|, |B|)$; {\color{blue} $(a,b) \in R$, $a \leq a'$,
  and $b \leq b'$ imply $(a',b') \in R$; so next part is redundant?}
and $(\bigvee_{i < \omega} a_i, \bigvee_{i < \omega} b_i) \in R$
whenever $(a_i)_{i < \omega}$ and $(b_i)_{i < \omega}$ are chains in
$A$ and $B$, respectively, such that $(a_i,b_i) \in R$ for all
$i$. The category $\oCPOR$ has as its morphisms from $(A,B,R)$ to
$(A',B',R')$ pairs $(f : A \to A',g : B \to B')$ of morphisms in
$\oCPO$ such that $(f\,a, g\,b) \in R'$ whenever $(a,b) \in R$. We
note that if $(f,g) : (A,B,R) \to (A',B',R')$ and $(a_i)_{i < \omega}$
and $(b_i)_{i < \omega}$ are chains in $A$ and $B$, respectively, then
$ (f \, (\bigvee_{i < \omega} a_i), g \, (\bigvee_{i < \omega} b_i)) =
(\bigvee_{i < \omega} (f a_i), \bigvee_{i < \omega} (g b_i)) \in R'$
necessarily holds.}  We write $R : \oCPOR(A,B)$ in place of $(A,B,R):
\oCPOR$ when convenient. If $R : \oCPOR(A,B)$ then we write $\pi_1 R$
and $\pi_2 R$ for the {\em domain} $A$ of $R$ and the {\em codomain}
$B$ of $R$, respectively. If $A : \oCPO$, then we write $\Eq_A =
(A,A,\Eq_{|A|})$ for the {\em equality relation} on $A$.

@@

The object interpretations of the types in our calculus will be
$\omega$-cocontinuous functors between categories of
$\omega$-cocontinuous functors on categories constructed from the
locally $\omega$-presentable category $\oCPO$.  We therefore begin
by recording some important facts about locally $\omega$-presentable
categories and functors on them, and verifying verify the properties
needed to interpret our syntax.

@@

\new{Perhaps have as preliminaries to entire paper.}

@@

A poset $\mcD = (D,\leq)$ is \emph{$\omega$-directed} if every
countable subset of $D$ has a supremum. When $\mcD$ is considered
as a category, we write $d \in \mcD$ to indicate that $d$ is an object
of $\mcD$ (i.e., $d \in D$). An {\em $\omega$-directed colimit} in a
category $\mcC$ is a colimit of a diagram $F : {\mathcal D} \to \mcC$,
where $\mathcal D$ is an $\omega$-directed poset. A category $\mcC$
is {\em $\omega$-cocomplete} if it has all $\omega$-directed
colimits; a {\em cocomplete} category is one that has all colimits.

If $\mcA$ and $\mcC$ are $\omega$-cocomplete, then a functor $F :
\mcA \to \mcC$ is {\em $\omega$-cocontinuous} if it preserves
$\omega$-directed colimits. We write $[\A,\C]_{\omega}$ for the
category of $\omega$-cocontinuous functors from $\A$ to $\C$, and
$\C^\A$ for the category of {\em all} functors from $\A$ to $\C$.
Since (co)limits of functors are computed pointwise, $C^\A$ has all
(co)limits that $\C$ has, and (co)limits of (co)continuous functors
are again (co)continuous. It follows that $[\A,\C]_{(\omega)}$ is
($\omega$-)(co)complete whenever $\C$ is. In particular, if $F :
[\C, \C]_{(\omega)}$ for a(n) ($\omega$-)cocomplete category $\C$,
then we can define $F^0 = \Id$, $F^{\alpha + 1} = F \circ F^\alpha$ at
successor ordinals, and $F^{\bigcup_{\beta < \alpha} \beta} =
\colim{\beta < \alpha} F^\beta$ at limit ordinals.

If $\mcA$ is locally small, then an object $A$ of $\mcA$ is {\em
  $\omega$-presentable} if the functor $\mathsf{Hom}_\mcA(A, -) :
\mcA \to \Set$ preserves $\omega$-directed colimits, i.e., if for
every $\omega$-directed poset $\mathcal D$ and every functor $F :
{\mathcal D} \to \A$, there is a canonical isomorphism $\colim{d \in
  \mcD}{\mathsf{Hom}_\mcA(A,Fd)} \simeq \mathsf{Hom}_{\mcA}(A,
\colim{d \in \mcD}{Fd})$. A locally small category $\mcA$ is {\em
  $\omega$-accessible} if it is $\omega$-cocomplete and has a set
$\mcA_0$ of $\omega$-presentable objects such that every object is
an $\omega$-directed colimit of objects in $\mcA_0$; a locally small
category is {\em locally $\omega$-presentable} if it is
$\omega$-accessible and cocomplete.

The category $\oCPO$ is locally presentable; its presentable objects
are precisely the finite sets.
%The category $\oCPO$ is locally $\omega$-presentable (but not
%locally finitely presentable); see Examples~1.18(2)
%of~\cite{ar94}). Its $\omega$-presentable objects are precisely the
%$\omega$CPOs cardinality less than $\omega$, i.e., the countable
%$\omega$CPOs.
In the next subsection we will interpret type variables in $\tvars^k
\cup \fvars^k$ as elements of $[\oCPO^k,\oCPO]_{\omega}$; the
following special cases of standard results (see, e.g.,~\cite{ar94})
will therefore be critical to deducing important properties of our
object interpretations of types:
\begin{prop}\label{prop:prelims}
\begin{enumerate}
\item If $\C_1,...,\C_n$ are locally $\omega$-presentable categories
  then so is $\C_1 \times ... \times \C_n$. Moreover, the presentable
  objects of $\C_1 \times ... \times \C_n$ are exactly the tuples of
  the form $(P_1,...,P_n)$, where, for each $i = 1,...,n$, the object
  $P_i$ is presentable in $\C_i$.
\item If $\mathcal{A}$ is $\omega$-accessible and $\mathcal{C}$ is
  $\lambda$-cocomplete, then the category $[\mathcal{A},
  \mathcal{C}]_{\omega}$ is naturally equivalent to the category
  $\mathcal{C}^{\mathcal{A}_0}$.
\item If $\mathcal{C}$ is locally $\omega$-presentable and
  $\mathcal{A}_0$ is essentially small, then
  $\mathcal{C}^{\mathcal{A}_0}$ is locally $\omega$-presentable.
\end{enumerate}
\end{prop}
\noindent
Together, the statements in Proposition~\ref{prop:prelims} give that
if $\mathcal{A}$ and $\mathcal{C}$ are locally $\omega$-presentable,
then $[\mathcal{A}, \mathcal{C}]_{\omega}$ is naturally equivalent
to $\mathcal{C}^{\mathcal{A}_0}$, and thus is
$\omega$-presentable. Thus, for all $k_1,...,k_n \in \nat^n$,
$[\mathcal{A}^{k_1}, \mathcal{C}]_{\omega} \times ... \times
[\mathcal{A}^{k_n}, \mathcal{C}]_{\omega}$ is locally
$\omega$-presentable, and therefore $[[\mathcal{A}^{k_1},
    \mathcal{C}]_{\omega} \times ... \times [\mathcal{A}^{k_n},
    \mathcal{C}]_{\omega}, \mathcal{C}]_{\omega}$ is as
well. Taking both $\A$ and $C$ to be $\oCPO$ --- as we will to ensure
that the fixpoints interpreting $\mu$-types in $\oCPO$ exist --- we
have
\begin{prop}\label{prop:nat-interps-sets}
  For all $k_1,...,k_n \in \nat^n$,
  \[[[\oCPO^{k_1}, \oCPO]_{\omega}
  \times ... \times [\oCPO^{k_n}, \oCPO]_{\omega},
  \oCPO]_{\omega}\] is locally $\omega$-presentable.
\end{prop}

@@

\begin{dfn}\label{def:set-env}
A {\em $\oCPO$ environment} maps each type variable in $\tvars^k \cup
\fvars^k$ to an element of $[\oCPO^k,\oCPO]_{\omega}$.  A morphism
$f : \rho \to \rho'$ for set environments $\rho$ and $\rho'$ with
$\rho|_\tvars = \rho'|_\tvars$ maps each type constructor variable
$\psi^k \in \tvars$ to the identity natural transformation on $\rho
\psi^k = \rho'\psi^k$ and each functorial variable $\phi^k \in \fvars$
to a natural transformation from the $k$-ary functor $\rho \phi^k$ on
$\oCPO$ to the $k$-ary functor $\rho' \phi^k$ on $\oCPO$.  Composition
of morphisms on set environments is given componentwise, with the
identity morphism mapping each set environment to itself. This gives a
category of set environments and morphisms between them, which we
denote $\oCPOenv$.
\end{dfn}

When convenient we identify a functor in $[\oCPO^0, \oCPO]_{\omega}$
with its value on $\ast$ and consider a $\oCPO$ environment to map a
type variable of arity $0$ to an $\omega$-cocontinuous functor from
$\oCPO^0$ to $\oCPO$, i.e., to an $\oCPO$. If $\Phi =
\{\phi_1^{k_1},...,\phi_n^{k_n}\}$ and $\ol K = \{K_1,...,K_n\}$ where
$K_i : [\oCPO^{k_i},\oCPO]_{\omega}$ for $i = 1,...,n$, then we
write either $\rho[\ol{\Phi := K}]$ or $\rho[\ol{\phi := K}]$ for the
$\oCPO$ environment $\rho'$ such that $\rho' \phi_i = K_i$ for $i =
1,...,n$ and $\rho' \phi = \rho \phi$ if $\phi \not \in \Phi$.  If
$\rho$ is an $\oCPO$ environment, we write $\Eq_\rho$ for the $\oCPOR$
environment (see Definition~\ref{def:reln-env}) such that $\Eq_\rho v
= \Eq_{\rho v}$ for every type variable $v$. The categories
$\oCPORT_k$ and relational interpretations appearing in the third
clause of Definition~\ref{def:set-sem} are given in full in
Section~\ref{sec:rel-interp}. We note that $\omega$-directed
colimits in $\oCPOenv$ are taken pointwise.
